---
title: "Pulling Franklin County ACS Data in R"
author: "Josh Baney"
date: "6/3/2020"
output:
  html_document:
    df_print: paged
    theme: paper
---

### Introduction

This notebook walks through the process for pulling and mapping Franklin County data from the American Community Survey using the Tidycensus Package.  This is simply a quick adaptation of the much more in depth tutorial located here:  <https://dlab-berkeley.github.io/Census-Data-in-R/Rcensus_data_maps-slides.html#1>


### Load R Packages

The following packages are needed:

- [tidycensus](https://walkerke.github.io/tidycensus) & [tigris](https://github.com/walkerke/tigris)

- [tidyverse](https://www.tidyverse.org/)
   
- [sf](https://r-spatial.github.io/sf/)

```{r, echo = TRUE, message = FALSE, warning=FALSE}
library(tidycensus)
library(tidyverse) 
library(tigris)
library(sf)
```
<p>&nbsp;</p>
### Install Census API Key

Use the tidycensus function **census_api_key** to make tidycensus use your key when it fetches data from the census.
In order to begin using the TidyCensus package a Census API Key must be acquired and installed on machine. Key can be aquired here: https://api.census.gov/data/key_signup.html 

```{r, eval = F}
# Install Census API Key on R Instance
# census_api_key('Insert Key Here', install = TRUE)
```
<p>&nbsp;</p>
### View all available ACS Variables\
```{r}
# Load All ACS Variables for 2018 (Most Recent Year) and Examine Structure
acs2018vars <- load_variables(year=2018, dataset = 'acs5', cache = T)
str(acs2018vars)
```
<p>&nbsp;</p>
### Load ACS Variables using the get_acs function

    C17002_002:  Estimated Pop. under 50% of Poverty Line
    C17002_003: Estimated Pop. between 50% and 99% of Poverty Line
    C17002_001: Estimated Pop., Serves as Denominator
```{r, message = FALSE, warning = FALSE, results='hide' }

fr_poor <- get_acs(geography = "tract",  
                   variables = c('C17002_002','C17002_003'), # poverty variables
                   year = 2018,          
                   state="OH",
                   summary_var = "C17002_001", # Est of num people - denom
                   county="Franklin",
                   geometry=T)

```

```{r}
head(fr_poor)
```
<p>&nbsp;</p>
### Aggregate Count and Percentage Below Poverty Line
```{r}
fr_poor_summed <- fr_poor %>% 
            filter(summary_est > 0) %>% # Remove Census Tracts with No Population
            select(GEOID, estimate, summary_est,geometry) %>%
            group_by(GEOID) %>% 
            summarise(count_below_pov = sum(estimate), # Aggregate Count Below Poverty Line
            total_pop = max(summary_est)) %>% # Use Max to Avoid Duplicating Total Pop
            mutate(pct = 100 * (count_below_pov / total_pop)) # Calculate % Below Poverty Line

head(fr_poor_summed)
```
<p>&nbsp;</p>
### Map Total Count Below Poverty Line 
```{r}
plot(fr_poor_summed['count_below_pov'], main = 'Count Below Poverty Line')
```
<p>&nbsp;</p>
### Map Percent Below Poverty Line
```{r}
plot(fr_poor_summed['pct'], main = 'Percent Below Poverty Line')

```
<p>&nbsp;</p>
### Create an interactive Map using TMAP Package

```{r}
library(tmap)
tmap_mode("view") # set mode to interactive
# Create Map Object
poverty_map <- tm_shape(fr_poor_summed) +
                  tm_polygons(col="pct", alpha = 0.8) 
```

```{r}
# View Map
poverty_map
```

OSU Students living off campus appear to be effecting poverty rates.  More info can be found here: https://www.census.gov/library/stories/2018/10/off-campus-college-students-poverty.html